---
title: "post MR operation"
author: "Lanlan Chen"
date: "6/14/2019"
output: html_document
---

```{r setup, include=FALSE,message=FALSE}
library(TwoSampleMR)
library(ggplot2)
library(cowplot)
library(diagram)
```

## Statistical method and programming for MR

```{r}
wald_ratio <- function(b_exp, b_out, se_exp, se_out, parameters)
{
	if(length(b_exp) > 1)
	{
		return(list(b=NA, se=NA, pval=NA, nsnp=NA))
	}
	b <- b_out / b_exp
	se <- se_out / abs(b_exp)
	# sqrt((segd^2/gp^2) + (gd^2/gp^4)*segp^2 - 2*(gd/gp^3)) #full delta method with cov set to 0
	pval <- pnorm(abs(b) / se, lower.tail=FALSE) * 2
	return(list(b=b, se=se, pval=pval, nsnp=1))
}

ivw <- function(b_exp, b_out, se_exp, se_out, parameters=default_parameters())
{
	if(sum(!is.na(b_exp) & !is.na(b_out) & !is.na(se_exp) & !is.na(se_out)) < 2)
	return(list(b=NA, se=NA, pval=NA, nsnp=NA))

	ivw.res <- summary(lm(b_out ~ -1 + b_exp, weights = 1/se_out^2))
	b <- ivw.res$coef["b_exp","Estimate"]
	se <- ivw.res$coef["b_exp","Std. Error"]/min(1,ivw.res$sigma) #sigma is the residual standard error
	pval <- 2 * pnorm(abs(b/se), low=FALSE)
	Q_df <- length(b_exp) - 1
	Q <- ivw.res$sigma^2 * Q_df
	Q_pval <- pchisq(Q, Q_df, low=FALSE)
	# from formula phi =  Q/DF rearranged to to Q = phi*DF, where phi is sigma^2
	# Q.ivw<-sum((1/(se_out/b_exp)^2)*(b_out/b_exp-ivw.reg.beta)^2)
	return(list(b = b, se = se, pval = pval, nsnp=length(b_exp), Q = Q, Q_df = Q_df, Q_pval = Q_pval))
}
```

```{r}
wald_ratio(0.3, 0.004, 0.6,0.3)
ivw(b_exp=c(0.3,0.5),se_exp=c(0.004,0.0005),b_out=c(0.6,0.8),se_out=c(0.3,0.04))
```

## How to plot MR

Now we have got the MR results from our high throughput analysis and the next step is to extract those significant phenotypes and visualize our results of interest

First, I will introduce you to plot MR results:
```{r, echo=TRUE}
setwd("C:\\Users\\11780\\Desktop\\pannap_res\\my-1190_my-6150_2_both_sexes") # to set the working directory
dat = read.table('harmonized_data.tsv',sep='\t',header=T) # to read preprocessed data for next analysis
head(dat,n=3) # to view the first 3 lines of data
dat$id.exposure = 'Daytime napping'
dat$id.outcome = 'Angina'
dat$exposure = 'Daytime napping'
dat$outcome ='Angina' # to change the name of exposure and outcome
head(dat,n=3) # to see the changed data
mr_method_list() # to look at all the methods available for TwoSampleMR package
```

Please remember to change the phenotype to your desired one and it is essential to next steps.

Here, we can get an overview of our original data. Then we are about to plot.

Please note that there are many methods available for performing MR and you can choose whatever you think is appropriate. However, I suggest using the three one I have mentioned above. 

### plot a MR scatter picture

```{r, echo=TRUE}
res = mr(dat, method_list=c('mr_ivw','mr_egger_regression','mr_weighted_median')) # to get the MR results
print(res)
p1 <- mr_scatter_plot(res,dat)[[1]]
print(p1)
```

### plot a MR forest picture

```{r, echo=TRUE}
res_single = mr_singlesnp(dat,all_method=c('mr_ivw','mr_egger_regression','mr_weighted_median')) ## to contruct the dataframe used in forest plot
p2 <- mr_forest_plot(res_single)[[1]]
print(p2)
```

### plot a MR leave-one-out sensitivity picture

```{r, echo=TRUE}
loop = mr_leaveoneout(dat)
p3 <- mr_leaveoneout_plot(loop)[[1]]
print(p3)
```

### plot a MR funnel picture

```{r, echo=TRUE}
p4 <- mr_funnel_plot(res_single)[[1]]
print(p4)
```

### to combine the 4 pictures into one 

```{r, echo=TRUE}
p <- plot_grid(p1,p2,p3,p4)
print(p)
```

'cowplot' is a powerful packages for combining pictures and plot_grid() is the most useful function in it.

## Plot a picture to demonstrate our discovery

```{r, echo=TRUE}
##to plot the causal relationship
#setwd("C:\\Users\\11780\\Documents\\R语言练习\\Diagram")
#pdf('pan_nap_mr.pdf',height=6.5, width=14)
par(mar=c(1,1,1,1))
openplotmat(main="")
plotpos <- coordinates(c(3,3,3,3))
straightarrow( from=plotpos[8,], to=plotpos[4,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[2,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[6,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[1,], lwd=3, lty=2, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[3,], lwd=3, lty=2, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[7,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[9,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[10,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[11,], lwd=3, arr.pos=0.5, arr.length=0.5)
straightarrow( from=plotpos[8,], to=plotpos[12,], lwd=3, arr.pos=0.5, arr.length=0.5)
textdiamond(plotpos[2,], 0.09, 0.05, lab=c('Lethargy', 'Tiredness'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textellipse(plotpos[1,], 0.09, 0.05, lab=c('Walking pace'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textellipse(plotpos[3,], 0.09, 0.05, lab=c('Walking frequency'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textround(plotpos[8,], 0.09, 0.05, lab=c('Daytime napping'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textparallel(plotpos[4,], 0.09, 0.05, lab=c('Diastolic','blood pressure'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textparallel(plotpos[6,], 0.09, 0.05, lab=c('Body fat','percentage'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textrect(plotpos[7,], 0.09, 0.05, lab=c('Hematocrit','percentage'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textrect(plotpos[9,], 0.09, 0.05, lab=c('Reticulocyte','count'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textrect(plotpos[10,], 0.09, 0.05, lab=c('Angina'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textrect(plotpos[11,], 0.09, 0.05, lab=c('Gastroesophageal','reflux'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
textrect(plotpos[12,], 0.09, 0.05, lab=c('Asthma'),
         box.col="gray", shadow.col="darkgray", shadow.size=0.005, cex=1.1, font=2)
#dev.off()
```

## Thank you for your listening!